---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tibblify

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/tibblify)](https://CRAN.R-project.org/package=tibblify)
[![Codecov test coverage](https://codecov.io/gh/mgirlich/tibblify/branch/master/graph/badge.svg)](https://codecov.io/gh/mgirlich/tibblify?branch=master)
[![R build status](https://github.com/mgirlich/tibblify/workflows/R-CMD-check/badge.svg)](https://github.com/mgirlich/tibblify/actions)
<!-- badges: end -->

The goal of tibblify is to provide an easy way of converting a nested list into a tibble.

## Installation

You can install the released version of tibblify from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("tibblify")
```

Or install the development version from GitHub with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("mgirlich/tibblify")
```


```{r child='vignettes/tibblify.Rmd'} 
```


## Code of Conduct

Please note that the tibblify project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
