---
title: "Overview supported structures"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview supported structures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tibblify)
```

## Supported input for `tibblify()`

The idea of `tibblify()` is to make it easier and more robust to convert lists
of lists into tibbles. This is a typical task after receiving API responses in
JSON format.
The following provides an overview which kind of R objects are supported and
the JSON they correspond to.


### Scalars

There are 4 basic types of scalars coming from JSON: boolean, integer, float, string

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px; width: 100%;"}

::: {}

```{json}
true
1
1.5
"a"
```

:::

::: {}

```{r results='hide'}
TRUE
1
1.5
"a"
```

:::

::::

Other R vectors without JSON equivalent are also supported as long as they:

* are a vector in the vctrs definition and
* have size one, i.e. `vctrs::vec_size(x)` is 1.

Examples are `Date` or `POSIXct`

In general a scalar can be parsed with `tib_scalar()`. There are some special
functions for common types:

* `tib_lgl()`
* `tib_int()`
* `tib_dbl()`
* `tib_chr()`
* `tib_date()`

can be parsed with

```{r eval=FALSE}
tib_scalar(ptype = vctrs::new_datetime())
```


### Vectors

A homogeneous JSON array is an array of scalar where each scalar has the same type.
In R they correspond to a `logical()`, `integer()`, `double()` or `character()`
vector:

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px; width: 100%;"}

::: {}

```{json}
[true, null, false]
[1, null, 3]
[1.5, null, 3.5]
["a", null, "c"]
```

:::

::: {}

```{r eval=FALSE}
c(TRUE, NA, FALSE)
c(1L, NA, 2L)
c(1.5, NA, 2.5)
c("a", NA, "c")
```

Or if parsed with `jsonlite::fromJSON(simplifyVector = FALSE)`

```{r eval=FALSE}
list(TRUE, NULL, FALSE)
list(1L, NULL, 2L)
list(1.5, NULL, 2.5)
list("a", NULL, "c")
```

:::

::::

They can be parsed with `tib_vector()`. As for scalars there are shortcuts for
some common types, e.g. `tib_lgl_vec()`.

#### Empty lists

A special case are empty lists `list()`. They might appear when parsing an empty
JSON array:

```{r error=TRUE}
x_json <- '[
  {"a": [1, 2]},
  {"a": []}
]'

x <- jsonlite::fromJSON(x_json, simplifyDataFrame = FALSE)
str(x)
```

By default they are not supported but produce an error:

```{r error=TRUE}
tibblify(x, tspec_df(tib_int_vec("a")))
```

Use `vector_allows_empty_list` argument of `tspec_*()` so that they are converted
to an empty vector instead:

```{r}
tibblify(x, tspec_df(tib_int_vec("a"), vector_allows_empty_list = TRUE))$a
```

#### Homogeneous R lists of scalars

When using `jsonlite::fromJSON(simplifyVector = FALSE)` to parse JSON to an R
object one does not get R vectors but lists of scalars:

```{r error=TRUE}
x_json <- '[
  {"a": [1, 2]},
  {"a": [1, 2, 3]}
]'

x <- jsonlite::fromJSON(x_json, simplifyVector = FALSE)
str(x)
```

By default they cannot be parsed with `tib_vector()`

```{r error=TRUE}
tibblify(x, tspec_df(tib_int_vec("a")))
```

Set the argument `input_form` to `"scalar_list"` to parse them:

```{r}
tibblify(x, tspec_df(tib_int_vec("a", input_form = "scalar_list")))$a
```


### Homogeneous JSON objects of scalars

To parse vector like structures which are encoded as JSON objects:

```{r}
x_json <- '[
  {"a": {"x": 1, "y": 2}},
  {"a": {"a": 1, "b": 2, "b": 3}}
]'

x <- jsonlite::fromJSON(x_json, simplifyVector = FALSE)
str(x)
```

set the argument `input_form` to `"object"`. To actually store the names use the `names_to` and `values_to` argument:

```{r}
spec <- tspec_df(
  tib_int_vec(
    "a",
    input_form = "object",
    names_to = "name",
    values_to = "value"
  )
)
tibblify(x, )$a
```

### Varying

Lists where elements do not have a common type but vary:

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px; width: 100%;"}

::: {}

```{json}
[1, "a", true]
```

:::

::: {}

```{r eval=FALSE}
list(1, "a", TRUE)
```

:::

::::

can be parsed with `tib_variant()`.


### 


### NULL

`tibblify()` supports four formats:

### 1. Object

The R equivalent to a JSON object is a named list where the names fulfill the
requirements of `vctrs::vec_as_names(repair = "check_unique")`.

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px; width: 100%;"}

::: {}

```{json}
{
  "a": 1,
  "b": true
}
```

:::

::: {}


```{r results='hide'}
x <- list(
  a = 1,
  b = TRUE
)
```

:::

::::

They can be parsed with `tspec_object()` to produce a list or with `tspec_row()`
to produce a 1 row tibble

```{r}
object_spec <- tspec_object(
  tib_int("a"),
  tib_lgl("b")
)

tibblify(x, object_spec)
```

```{r}
row_spec <- tspec_row(
  tib_int("a"),
  tib_lgl("b")
)

tibblify(x, row_spec)
```


### 2.  List of homogeneous objects

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px; width: 100%;"}

::: {}

```{json}
[
  {"a": 1, "b": true},
  {"b": 2, "b": false}
]
```

:::

::: {}

```{r results='hide'}
x <- list(
  list(a = 1, b = TRUE),
  list(a = 2, b = FALSE)
)
```

:::

::::

They are parsed with `tspec_df()`

```{r}
df_spec <- tspec_df(
  tib_int("a"),
  tib_lgl("b")
)

tibblify(x, df_spec)
```

### 3.  Object of homogeneous objects

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px; width: 100%;"}

::: {}

```{json}
{
  "object1": {"a": 1, "b": true},
  "object2": {"b": 2, "b": false}
}
```

:::

::: {}

```{r results='hide'}
x <- list(
  object1 = list(a = 1, b = TRUE),
  object2 = list(a = 2, b = FALSE)
)
```

:::

::::

They are also parsed with `tspec_df()` but you can parse the names into an extra
column via the `.names_to` argument:

```{r}
df_spec <- tspec_df(
  tib_int("a"),
  tib_lgl("b"),
  .names_to = "name"
)

tibblify(x, df_spec)
```


### 4.  Object in column major format

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px; width: 100%;"}

::: {}

```{json}
{
  "a": [1, 2],
  "b": [true, false]
}
```

:::

::: {}

```{r results='hide'}
x <- list(
  a = c(1, 2),
  b = c(TRUE, FALSE)
)
```

:::

::::

They are also parsed with `tspec_df()` but you need to use `.input_form = "colmajor"`:

```{r}
df_spec <- tspec_df(
  tib_int("a"),
  tib_lgl("b"),
  .input_form = "colmajor"
)

tibblify(x, df_spec)
```
