---
title: "tibblify"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tibblify}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
```

## Usage

Let's convert the Leaders dataset into a tibble. It is a list with one element per character:

```{r usage_example}
library(tibblify)

str(politicians[1])
```

We can let `tibblify()` automatically recognize the structure of the list and find an appropriate presentation as a tibble:

```{r usage_leaders}
politicians_tibble <- tibblify(politicians)
politicians_tibble
```

The `parents` column is a tibble with the columns `mother` and `father` because in the original list `leader1` the field `parents` is a named list.

```{r usage_tibble_column}
politicians_tibble$parents
```

and the `spouses` column is a [`list_of`](https://vctrs.r-lib.org/reference/list_of.html) character because the `spouses` field is a list and all elements are characters

```{r usage_list_of_column}
politicians_tibble$spouses
```


## Specification

In the above example we used `tibblify()` without any further specification on how to convert the list into a tibble. This is quite useful in an interactive session but often you want to provide a specification yourself. Some of the reasons are:

* to ensure type and shape stability of the resulting tibble in automated scripts.
* to use a different type of a column.
* to use different names.
* to parse only a subset of columns.
* to specify what happens if a value is missing.

First, we use `get_spec()` to view the specification used to convert our list to a tibble:

```{r specification}
get_spec(politicians_tibble)
```

A specification always starts with a call to `lcols()` (similar to [`readr::cols()`](https://readr.tidyverse.org/reference/cols.html)). Then you specify the columns you want with name-value pairs. The name is the name of the resulting column and the value is a specification created with one of the `lcol_*()` functions.


## Path

The first argument to `lcol_*()` is always a `path` which describes where to find the element. The syntax is the same as in `purrr::map()` used to extract fields. Some examples

```{r path_examples}
leader <- politicians[[1]]

# get the element `id`
path <- c("id")
leader[["id"]]

# get the element `father` in the element `parents`
path <- c("parents", "father")
leader[["parents"]][["mother"]]

# get the first element in the element `spouses`
path <- list("spouses", 1)
leader[["spouses"]][[1]]
```


## Vector Columns

A couple of typical vector types have a predefined extractor:

* `lcol_chr()`: create a character column.
* `lcol_lgl()`: create a logical column.
* `lcol_int()`: create an integer column.
* `lcol_dbl()`: create a double column.
* `lcol_dat()`: create a date column.
* `lcol_dtt()`: create a datetime column.

See [parsing other types] to create a column of your own prototype.

```{r vector_cols}
tibblify(
  politicians,
  lcols(
    lcol_int("id"),
    lcol_chr("name"),
    `family name` = lcol_chr("surname")
  )
)
```


## Missing Elements

If an element doesn't exist an error is thrown as in `purrr::chuck()`. To use a default value instead of throwing an error use the `.default` argument. The `.default` value is also used in case the element at the path is empty:

```{r missing_elements, error=TRUE}
list_default <- list(
  list(a = 1),
  list(a = NULL),
  list(a = integer()),
  list()
)

tibblify(
  list_default,
  lcols(lcol_int("a"))
)

tibblify(
  list_default,
  lcols(lcol_int("a", .default = 0))
)
```


## Parser

When the cast is not possible with `vctrs::vec_cast()` you can use the `.parser` argument to supply a custom parser. It is passed to `rlang::as_function()` so you can use a function or a formula. A typical use case are dates stored as strings.

```{r custom_parser}
tibblify(
  politicians,
  lcols(
    lcol_chr("surname"),
    lcol_dat("dob", .parser = ~ as.Date(.x, format = "%Y-%m-%d"))
  )
)
```


## List and List Of Columns

A `list_of` is a list where each element in the list has the same prototype. It is useful when you have fields with more than one element as in the `spouses` field.

```{r}
spouses_tbl <- tibblify(
  politicians,
  lcols(
    lcol_chr("surname"),
    lcol_lst_of("spouses", .ptype = character(), .parser = ~vctrs::vec_c(!!!.x, .ptype = character()))
  )
)

spouses_tbl$spouses
```

You can use [`tidyr::unnest()`](https://tidyr.tidyverse.org/reference/nest.html) or [`tidyr::unnest_longer()`](https://tidyr.tidyverse.org/reference/hoist.html) to flatten these columns to regular columns.

A list column is used when you have a field with mixed elements.


## Guess and Skip

Analogue to `readr::col_guess()` and `readr::col_skip()` you can specify that you want to guess the column type with `lcol_guess()` respectively skip a field with `lcol_skip()`. Skipping a column can be useful when you set a default column type or you want to make clear that you know about the field and intentionally skip it.

Guessing a column is useful in interactive sessions but you shouldn't rely on it in automated scripts.


## Tibble Columns and List Of Tibble Columns

If a field contains is a named list where each element has length 1 or 0 the field is converted to a tibble column. This is for example the case for the `parents` field:

```{r}
leaders_tibble <- tibblify(
  politicians,
  lcols(
    lcol_chr("surname"),
    lcol_guess("parents")
  )
)

leaders_tibble
```

Tibble columns are a relatively new concept in the tidyverse. You can unpack a tibble column into regular columns with `tidyr::unpack()`.



## Parsing other types

`tibblify` provides shortcuts for a couple of common types. To parse a vector or record type without a parser use `lcol_vec()`. Let's say you have a list with `difftimes`

```{r}
now <- Sys.time()
past <- now - c(100, 200)

x <- list(
  list(timediff = now - past[1]),
  list(timediff = now - past[2])
)

x
```

You need to define a prototype

```{r}
ptype <- as.difftime(0, units = "secs")
ptype
```

and then use it in `lcol_vec()`

```{r}
tibblify(
  x,
  lcols(
    lcol_vec("timediff", ptype = ptype)
  )
)
```


## Default Column Type

You can use the `.default` argument of `lcols()` to define a parser used for all unspecified fields.

```{r}
tibblify(
  politicians,
  lcols(
    lcol_chr("name"),
    lcol_chr("surname"),
    .default = lcol_lst(path = zap(), .default = NULL)
  )
)
```

